---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tidyvader

<!-- badges: start -->
<!-- badges: end -->

A fast, clear, and tidy implementation of the rule-based sentiment analysis algorithm VADER (Valence Aware Dictionary and Sentiment Reasoner). 

* **It's fast** because the code has been re-written from scratch and uses C++ for the core algorithm: in benchmarking it's over 500 times faster than the competing R package. 
* **It's clear** because it tries to make the rule-based algorithm and the dictionaries simple to read so that users can inspect and judge them. 
* **It's tidy** because it tries to follow tidy design principles and works well with the %>% pipe.

## Under Development

*Please note* that this package (and this documentation) is under active development. At present it's pretty well tested and functional, but this is not yet on CRAN and things may change. Expect more/better documentation as soon as time allows.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("chris31415926535/tidyvader")
```

## What is VADER?


## Example

This example shows how to send sentences in a dataframe through `vader()`. It also shows how punctuation, capitalization, and emojis all work together to affect a sentence's compound score.

```{r example, warning=FALSE, message=FALSE}
library(tidyvader)
library(tibble)
library(magrittr)

# set up a tibble with some sentences
texts <- tibble(sentences = c("I feel happy today.",
                              "I feel happy today!",
                              "I feel HAPPY today!",
                              "I feel HAPPY today! :)",
                              "But somtimes I feel sad :("))

# pipe the data to tidyvader::vader() and specify the column with text 
texts %>%
  tidyvader::vader(sentences) %>%
  knitr::kable()
```
